FINE SCALE REGISTRATION OF SIDEWALKS AND OTHER RIBBON-LIKE FEATURES FROM AERIAL IMAGERY

### 2018-11-06 Chapters/3.Hypothesis/Paper.tex
Corrections and updates to notation

\chapter{Detail Experiments on \ac{DPST}}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.8\textwidth]{Figures/diagram6.png}
\caption[Framework Overview]{Our approach adopted to predict precise boundaries for ribbon-like features. Where black line indicated a sidewalk's geometric information, and red lines shown our result for the sidewalk boundaries. From (a) to (f), the 6 step approach introduce input map (a), initial trajectory (b), ribbon image generation (c), density Estimation for pixels(d), the result from our \ac{DP} solution(e), the output result by reshape sidewalk(f).}
\label{fig:Apparatus}
\end{center}
\end{figure}

@@

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/arizon_center.png}
    \caption[Demonstration on Arizona Center]{Detail demonstration on part of Arizona center, all sidewalks' \ac{GIS} information shows in Orange.}
    \label{fig:arizon_center}
\end{figure}

Before applying deep learning network to determine sidewalk's location, it's necessary to hand marking all sidewalk from imagery or extract the geometric information of each sidewalk from resources like \ac{OSM}. 
It's challenging since the geometric accuracy was not precise, which could cause performance issue since we rely on this information to segment the sidewalk part from maps for the next step. 
To ensure data accuracy, as shows in figure \ref{fig:arizon_center}, we spent days marking all sidewalks and their precise boundaries in Arizona Central for initial experiments. 
We can locate each sidewalk by using the geometric information and saved them into separate files. 
We also calculate the average sidewalk width from boundaries to decide reasonable window size in the following step.

@@

\section{Density Estimation}
We aim to find a new trajectory $\OutputTrajectory{[0\dots m]}$ and radius $\OutputRadius{[0\dots m]}$ that are most likely given an input image. We first must estimate the probability of observing a color within the ribbon.  
Given an input image $\RibbonImage{}$, let $\Pr(\text{pos}|\phi,\theta)$ denote the probability that a pixel is a part of a ribbon given an observed color $\phi$ and set of learnable parameters $\theta$. 
We aim to estimate parameters $\theta$ that maximize the likelihood of our \textit{initial} path.

@@

We use the EM-GMM algorithm to find a max-likelihood estimate of $\theta$ as a pair of Gaussian mixtures so that $\Pr(\phi|\text{pos}, \theta)$ is a mixture of four multivariate Gaussian functions and $\Pr(\phi|\text{neg}, \theta)$ is a mixture of eight; then $\Pr(\text{pos}|\phi, \theta)$ can easily be found using Bayes theorem as 
$$\Pr(\text{pos}|\phi, \theta)= \frac{\Pr(\phi|\text{pos}, \theta)}{\Pr(\phi|\text{pos}, \theta)+\Pr(\phi|\text{neg}, \theta)}.$$

@@

The resulting probabilities are illustrated in \figref{fig:GMM_result_2}; we observe that even a coarse initial trajectory is usually enough to learn a more detailed, but noisy, estimate for the foreground object shape.
However colors alone are not able to discriminate between occlusion and camouflage regions (such as the driveway in \figref{fig:GMM_result}). 
For rotational convenience we let $p[i,j]=-\lg \Pr(\text{pos}|\RibbonImage{}[i,j], \theta)$ and $q[i,j]=-\lg \Pr(\text{neg}|\RibbonImage{}[i,j], \theta)$; then the \ac{NLL} of a ribbon is proportional to a sum of $p[i,j]$ for all $i,j$ within the ribbon and $q[i,j]$ for all $i,j$ outside the ribbon. 
The aim of our \ac{DP} solution is to simultaneously optimize the probability based on color and shape of a ribbon-like feature.

As shows in figure \ref{fig:GMM_Sample_2}, we applied a density estimation function to find the probability for each pixel in the ribbon image. 
For foreground density, we used grey-scale to determine the probability likely-hood for the sidewalk-feature, where white is more likely to be sidewalk and black is less likely. 
The background density is the opposite, which made black is more likely to be sidewalk and white is not. 
We use the log likely-hood function to calculate the foreground and background density and estimate density on pixels.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/GMM_SAMPLE2.png}
    \caption[Density Estimation on Sample Sidewalk]{A sample figure with density estimation applied on figure \ref{fig:Sample_Sidewalk_2}. Where white indicates pixels are more likely to be a sidewalk feature and black indicates pixels are more likely to be a non-sidewalk feature. We separate it into 4 parts from column 1 to 4.}
    \label{fig:GMM_Sample_2}
\end{figure}

@@

We take $\RibbonImage{}$ as an input ribbon image with $m$ rows and $n=2\MaxDistance+1$ columns; $\RibbonImage{}{[i,j]}$ is a pixel in the image. 
We have an estimate $\Pr(\text{pos}|\RibbonImage{}_{[i,j]}, \theta)$ which we abbreviate as $p_{[i,j]})$. 
In addition there is some small set of allowed perturbations $\delta=\{(\delta_j, \delta_k)\}$ where $\delta_j$ is a change in direction and $\delta_k$ is a change in the radius of the ribbon. 
Each perturbation has a probability $\Pr(\delta_j, \delta_k)$ that controls the shape of the curve; in our experiments we use $\delta_j\in\{-1,0,1\}, \delta_k\in\{-1, 0,1\}$ in order to ensure that results are connected, and  $\Pr(\delta_j, \delta_k)=\Pr(|\delta_j|)\Pr(\delta_k)$ is two independent multinomial distributions which require only three parameters due to symmetry and because probabilities sum to unity. 
We can interpret $c_\mathit{bend}=-\lg \Pr(\delta_j=\pm1)$ as a penalty for bending the curve, and $c_\mathit{shrink}=-\lg \Pr(\delta_k=-1)$ as a penalty for shrinking the radius, or $c_\mathit{expand} = -\lg \Pr(\delta_k=1)$ as a penalty for expanding the radius of the curve.

@@

\begin{codebox}
\Procname{$\proc{Segment-Ribbon}(\RibbonImage{})$} \label{alg:segmet-ribbin}
\li $p[i,j]\gets -\lg \Pr(\RibbonImage{}[i,j])/\MaxDistance{}$, 
          \hspace{2ex} $q[i,j] \gets -\lg \Pr(1-\RibbonImage{}[i,j])/\MaxDistance{}$
\li $C{[0\twodots{}m, 0\twodots{}n, 0\twodots{}n]} \gets \infty$
\li \For{$i \gets 0 \To m$} \Do                                                \label{li:loop-slices}
\li     \For $k \gets 0 \To \lfloor n/2 \rfloor $ \Do                          \label{li:loop-radius}
\li         $s = \displaystyle{\sum_{j=0}^{2k} p[i,j] +\sum_{j=2k+1}^n  q[i, j] }$
\li         \For $j \gets k \To n-k$ \Do  \label{li:loop-center}
\li              \If $i > 0$  \Do                                              \label{li:if-has-pred}
\li              $v \gets s {+}\displaystyle{\min_{\delta_j, \delta_k}\left(  
                              \begin{aligned}
                              C[i{-}1,j{+}\delta_j,k{+}\delta_k] \hspace{3ex}\\
                              \hfill{}-\lg \Pr(\delta_j, \delta_k)
                              \end{aligned}
                            \right)}$  
                               \label{li:min-from-pred}
\li              \Else $v\gets s$
                 \End
\li              $C{[i, j, k]} \gets v$                                       \label{li:dp-store}
\li              $s \gets s{+}q[i,j{-}k]{-}p[i, j{-}k]{+}p[i, j{+}k]{-}q[i,j{+}k]$ \label{li:update-s}
           \End
       \End
    \End
\li $j^*, k^* \gets \displaystyle{\arg\min_{j,k} C{[m, j, k]}}$
\li \Return $C, j^*, k^*$
\end{codebox}

The \proc{Backtrack-Ribbon} algorithm selects the least-costly (most probable) ribbon and it is copied into to $T$ and $R$; so that the most likely path has scale-space coordinates $(i, T[i], R[i])$ for $i=0\twodots{}m$.

\begin{codebox}
\Procname{$\proc{Backtrack-Ribbon}(C, j^*, k^*)$} \label{alg:backtrack-ribbon}
\li $T{[0{\twodots}m]} \gets 0$, \quad $R{[0{\twodots}m]} \gets 0$             
\li $T{[m]}, R{[m]} \gets j^*, k^*$ \label{li:backtrack-best-last}
\li \For $i \gets m-1 \Downto 0$ \Do
\li      $\delta_j, \delta_k = \displaystyle{\arg\min_{\delta_j,\delta_k}C{[i,j{-}\delta_j,k{-}\delta_l]}{-}\lg \Pr(\delta_j,\delta_k)}$
\li      $T{[i]}, R[i] \gets T{[i+1]}-\delta_j, R{[i+1]}-\delta_k$                                                 \label{li:backtack-choose-pred}
    \End
\li \Return $T, R, C_{[m, j^*, k^*]}$
\end{codebox}

@@

We use our dynamic programming algorithm on the ribbon image to find the max score for each pixel that had better probability to be sidewalk feature that we mentioned in Chapter 4. 
More importantly, we introduce a penalty control function that makes each horizontal movement in the edges or center cost more than vertical movement. Basically, for each line of pixels, if they think the edges or center should shift more horizontally than one pixel compares to the last line, we score them with more penalty.

>>>>

\chapter{Detail Experiments on \ac{DPRST}}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[width=0.8\textwidth]{Figures/diagram6.png}
\caption[Framework Overview]{Our approach adopted to predict precise boundaries for ribbon-like features. Where black line indicated a sidewalk's geometric information, and red lines shown our result for the sidewalk boundaries.
 From (a) to (f), the 6 step approach introduce 
(a) input map, 
(b)~initial trajectory, 
(c)~ribbon image generation, 
(d)~density Estimation for pixels, 
(e)~the result from our \ac{DP} solution, 
(f)~the output result.}
\label{fig:Apparatus}
\end{center}
\end{figure}

@@

\begin{figure}[H]
\centering 
\includegraphics[width=\textwidth]{Figures/arizon_center.png}
 \caption[Demonstration on Arizona Center]{
Ground-truth data for a part of Pheonix, Arizona. All sidewalk \ac{GIS}
information is shown in Orange.}
\label{fig:arizon_center}
\end{figure}

It's challenging since the geometric accuracy was not precise,
 which could cause performance issue since we rely on this 
 information to segment the sidewalk part from maps for the next step. 
To ensure data accuracy, as shown in figure \ref{fig:arizon_center},
 we spent days marking all sidewalks and their precise boundaries 
 in Arizona Central for initial experiments. 
We can locate each sidewalk by using the geometric information 
and saved them into separate files.

@@

\section{Density Estimation}\label{sec:density-estimation}
We aim to find a new trajectory $\OutputTrajectory{}[0\dots m]$ and radius $\OutputRadius{}[0\dots m]$ that are most likely given an input image. We first must estimate the probability of observing a color within the ribbon.  
Given an input image $\RibbonImage{}$, let $\Pr(R|\Color{},\theta)$ denote the probability that a pixel is a part of a ribbon given an observed color $\Color{}$ and set of learnable parameters $\theta$. 
We aim to estimate parameters $\theta$ that maximize the likelihood of our \textit{initial} path.

@@

We use the EM-GMM algorithm to find a max-likelihood estimate of $\theta$ as a pair of Gaussian mixtures so that $\Pr(\Color{}|R, \theta)$ is a mixture of four multivariate Gaussian functions and $\Pr(\Color{}|\neg R, \theta)$ is a mixture of eight; then $\Pr(R|\Color{}, \theta)$ can easily be found using Bayes theorem as 
$$\Pr(R|\Color{}, \theta)= \frac{\Pr(\Color{}|R, \theta)}{\Pr(\Color{}|R, \theta)+\Pr(\Color{}|\neg R, \theta)}.$$

@@

The resulting probabilities are illustrated in \figref{fig:GMM_result_2}; we observe that even a
coarse initial trajectory is usually enough to learn a more detailed, but noisy, estimate for the
foreground object shape. However colors alone are not able to discriminate between occlusion and
camouflage regions (such as the driveway in \figref{fig:GMM_result}).
Then the \ac{NLL} of a ribbon is proportional to a sum of $-\ln \Pr(R|\RibbonImage{}[i,j])$
for all $i,j$ within the ribbon and $-\ln \Pr(\neg R| \RibbonImage{}[i,j])$ for all $i,j$ outside the ribbon.
 The aim of our
\ac{DP} solution is to simultaneously optimize the probability based on color and shape of a
ribbon-like feature.

As shown in figure \ref{fig:GMM_Sample_2}, we applied a density estimation function to find the probability 
$\Pr(R|\RibbonImage[i,j])$ for each pixel in the ribbon image. 
For foreground density, we show the the probability likely-hood for 
the sidewalk-feature as gray-scale values, where white is more likely to be sidewalk and black is less likely. 
%We use the log likelihood function to calculate the foreground and background density and estimate density on pixels.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{Figures/GMM_SAMPLE2.png}
    \caption[Density Estimation on Sample Sidewalk]{
    A sample figure with density estimation applied on figure \ref{fig:Sample_Sidewalk_2}. Where
    white indicates pixels are more likely to be a sidewalk feature and black indicates pixels are
    more likely to be a non-sidewalk feature. We separate it into 4 parts from column 1 to 4.}
    \label{fig:GMM_Sample_2}
\end{figure}

@@

We take $\RibbonImage{}$ as an input ribbon image with $m$ rows and $n=2\MaxDistance+1$ columns; $\RibbonImage{}{[i,j]}$ is a pixel in the image. 
We have an estimate $\Pr(R|\RibbonImage{}_{[i,j]}, \theta)$ which we abbreviate as $p_{[i,j]}$. 
In addition there is some small set of allowed perturbations $\delta=\{(\delta_j, \delta_k)\}$ where $\delta_j$ is a change in direction and $\delta_k$ is a change in the radius of the ribbon. 
Each perturbation has a probability $\Pr(\delta_j, \delta_k)$ that controls the shape of the curve; in our experiments we use $\delta_j\in\{-1,0,1\}, \delta_k\in\{-1, 0,1\}$ in order to ensure that results are connected, and  $\Pr(\delta_j, \delta_k)=\Pr(|\delta_j|)\Pr(\delta_k)$ is two independent multinomial distributions which require only three parameters due to symmetry and because probabilities sum to unity. 
We can interpret $c_\mathit{bend}=-\lg \Pr(\delta_j=\pm1)$ as a penalty for bending the curve, and $c_\mathit{shrink}=-\lg \Pr(\delta_k=-1)$ as a penalty for shrinking the radius, or $c_\mathit{expand} = -\lg \Pr(\delta_k=1)$ as a penalty for expanding the radius of the curve.

@@

\begin{codebox}
\Procname{$\proc{Segment-Ribbon}(\RibbonImage{})$} \label{alg:segmet-ribbin}
\li $p[i,j]\gets -\lg \Pr(\RibbonImage{}[i,j])/\MaxDistance{}$, 
          \hspace{2ex} $q[i,j] \gets -\lg \Pr(1-\RibbonImage{}[i,j])/\MaxDistance{}$
\li $C{[0\twodots{}m, 0\twodots{}n, 0\twodots{}n]} \gets \infty$
\li \For{$i \gets 0 \To m$} \Do                                                \label{li:loop-slices}
\li     \For $k \gets 0 \To \lfloor n/2 \rfloor $ \Do                          \label{li:loop-radius}
\li         $s = \displaystyle{\sum_{j=0}^{2k} p[i,j] +\sum_{j=2k+1}^n  q[i, j] }$
\li         \For $j \gets k \To n-k$ \Do  \label{li:loop-center}
\li              \If $i > 0$  \Do                                              \label{li:if-has-pred}
\li              $v \gets s {+}\displaystyle{\min_{\delta_j, \delta_k}\left(  
                              C[i{-}1,j{+}\delta_j,k{+}\delta_k] -\lg \Pr(\delta_j, \delta_k)
                            \right)}$  
                               \label{li:min-from-pred}
\li              \Else $v\gets s$
                 \End
\li              $C{[i, j, k]} \gets v$                                       \label{li:dp-store}
\li              $s \gets s{+}q[i,j{-}k]{-}p[i, j{-}k]{+}p[i, j{+}k]{-}q[i,j{+}k]$ \label{li:update-s}
           \End
       \End
    \End
\li $j^*, k^* \gets \displaystyle{\arg\min_{j,k} C{[m, j, k]}}$
\li \Return $C, j^*, k^*$
\end{codebox}

The \proc{Backtrack-Ribbon} algorithm selects the least-costly (most probable) ribbon and it is
copied into to \OutputTrajectory{} and \OutputRadius{};
so that the most likely path has scale-space coordinates $(i, \OutputTrajectory{}[i],
\OutputRadius{}[i])$ for $i=0\twodots{}m$.

\begin{codebox}
\Procname{$\proc{Backtrack-Ribbon}(C, j^*, k^*)$} \label{alg:backtrack-ribbon}
\li $\OutputTrajectory{}{[0{\twodots}m]} \gets 0$, \quad $\OutputRadius{}[0{\twodots}m] \gets 0$             
\li $\OutputTrajectory{}{[m]}, \OutputRadius{}{[m]} \gets j^*, k^*$ \label{li:backtrack-best-last}
\li \For $i \gets m-1 \Downto 0$ \Do
\li      $\delta_j, \delta_k = \displaystyle{\arg\min_{\delta_j,\delta_k}C{[i,j{-}\delta_j,k{-}\delta_l]}{-}\lg \Pr(\delta_j,\delta_k)}$
\li      $\OutputTrajectory{}{[i]}, \OutputRadius{}[i] \gets \OutputTrajectory{}{[i+1]}-\delta_j, R{[i+1]}-\delta_k$                                                 \label{li:backtack-choose-pred}
    \End
\li \Return $\OutputTrajectory{}, \OutputRadius{}, C_{[m, j^*, k^*]}$
\end{codebox}

@@

%We use our dynamic programming algorithm on the ribbon image to find the max score for each pixel that had better probability to be sidewalk feature that we mentioned in Chapter 4. 
%More importantly, we introduce a penalty control function that makes each horizontal movement in the edges or center cost more than vertical movement. Basically, for each line of pixels, if they think the edges or center should shift more horizontally than one pixel compares to the last line, we score them with more penalty.
